A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-10-25 LyapDraft.tex
made several changes

We briefly mention the appearance of the $e^{-\alpha_t}$ term in~\eqref{Eq:E} and~\eqref{eq:general-sc-lyap}. In~\cite{Acceleration2}, this scaling was studied at length. In particular, it was shown that the family of Lagrangian functionals which generates~\eqref{Eq:EL} is invariant under the action of time reparameterization, which explain the appearance of the $e^{-\alpha_t}$ in front of $\dot X_t$. By contrast, the mirror descent dynamic, which is a gradient flow, is not invariant under time-reparameterization. This invariance property might suggest why momentum family is more natural for the purposes of optimization (that is, it might help us understand why the discrete-time implementations of the family of momentum dynamics achieves systematically better rates of convergence than the family of gradient flows).

@@

\begin{tabular}{|c|c|c|c|}

@@

@@

\subsection{Accelerated Gradient Descent}

@@

&= \left(1 - \frac{\alpha_k}{A_{k+1}} \right)\left(f(x_k) +\frac{1}{A_k} D_h(x, z_k)\right) + \frac{\alpha_k}{A_{k+1}} f_k(x)

@@

%
%
%
\subsection{Conditional Gradient Method} 
\paragraph{Continuous time Estimate Sequence}
For the conditional gradient method, the algorithm simply needed to ensure,
\begin{align}\label{Eq:InEq4}
0 &\leq \int_0^t\frac{d}{ds}\left\{e^{\beta_s}\right\}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle] -  \frac{d}{ds}\left\{ e^{ \beta_s}  f(X_s)\right\}\,ds
\end{align}
for our Lyapunov analysis to go through. Note, we can write~\eqref{Eq:InEq4} as, 
\begin{align*}
e^{ \beta_t}  f(X_t) \leq \int_0^t\frac{d}{ds}\left\{e^{\beta_s}\right\}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle]ds   + e^{\beta_0} f(X_0).
\end{align*}
From the derivation of the Lypaynov function, we can conclude
\begin{align*}
e^{\beta_t} f(X_t) \leq  (e^{\beta_t} - e^{\beta_0}) f(x) + e^{\beta_0}f(X_0).
\end{align*}
which shows that $\{f(X_t),e^{\beta_t}\}$ is a continuous-time estimate sequence for this method. In Appendix~\ref{App:EstSeq}, we show the equivalence in discrete-time between estimate sequences and our Lyapunov framework in this setting.
%

@@

  \clearpage
\bibliographystyle{plain}
\bibliography{refs.bib}
\clearpage

>>>>

We briefly mention the appearance of the $e^{-\alpha_t}$ term in~\eqref{Eq:E} and~\eqref{eq:general-sc-lyap}. This unusual given most Lyapunov functions are time-invariant. In~\cite{Acceleration}, this scaling was studied at length. In particular, it was shown that the family of Lagrangian functionals which generates~\eqref{Eq:EL} is invariant under the action of time reparameterization. By contrast, the mirror descent dynamic, which is a gradient flow, is not invariant under time-reparameterization. %Under the ideal scaling \eqref{Eq:IdeSca}, time reparameterization 
This invariance property explains why the Lyapunov function remains valid for the momentum dynamic~\eqref{Eq:E} which parameterized by $\alpha_t$ and $\beta_t$. 
%in these reparameterizations of time, however, you always have a valid lyapunov function
%This invariance property might suggest why momentum family is more natural for the purposes of optimization (that is, it might help us understand why the discrete-time implementations of the family of momentum dynamics achieves systematically better rates of convergence than the family of gradient flows).

@@

\begin{table}
\begin{tabular}{|c|c|c|c|}%\label{Eq:Table}

@@

\caption{How the Estimate Sequence is defined for the various algorithms}
\label{table:Table}
\end{table}

@@

\subsection{Equivalence between Estimate Sequences and Lyapunov Functions}
In this section, we show how one can show that estimate sequences are equivalent to our Lyapunov framework in both continuous and discrete time. 
The continuous time view shows that the errors in both the Lyapunov function and estimate sequences are due to discretization errors. We provide a sketch for how this works for most of these methods leaving some of the details for Appendix~\ref{App:EstSeq}.
\subsubsection{Accelerated Gradient Descent}

@@

&\overset{\text{Tab}~\ref{table:Table}}{=} \left(1 - \frac{\alpha_k}{A_{k+1}} \right)\left(f(x_k) +\frac{1}{A_k} D_h(x, z_k)\right) + \frac{\alpha_k}{A_{k+1}} f_k(x)

@@

%
%
%
\subsubsection{Conditional Gradient Method} 
\paragraph{Continuous time Estimate Sequence}
For the conditional gradient method, the algorithm simply needed to ensure,
\begin{align}\label{Eq:InEq4}
0 &\leq \int_0^t\frac{d}{ds}\left\{e^{\beta_s}\right\}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle] -  \frac{d}{ds}\left\{ e^{ \beta_s}  f(X_s)\right\}\,ds
\end{align}
for our Lyapunov analysis to go through. Note, we can write~\eqref{Eq:InEq4} as, 
\begin{align*}
e^{ \beta_t}  f(X_t) \leq \int_0^t\frac{d}{ds}\left\{e^{\beta_s}\right\}  [f(X_s)+ \langle \nabla f(X_s), x - X_s \rangle]ds   + e^{\beta_0} f(X_0).
\end{align*}
From the derivation of the Lypaynov function, we can conclude
\begin{align*}
e^{\beta_t} f(X_t) \leq  (e^{\beta_t} - e^{\beta_0}) f(x) + e^{\beta_0}f(X_0).
\end{align*}
which shows that $\{f(X_t),e^{\beta_t}\}$ is a continuous-time estimate sequence for this method. In Appendix~\ref{App:EstSeq}, we show the equivalence in discrete-time between estimate sequences and our Lyapunov framework in this setting.
%

@@

\bibliographystyle{plain}
\bibliography{refs.bib}
